---
title: "Population, GDP, Year"
author: "Tom Takeuchi"
date: "Spring 2022"
output: html_document
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding,
      output_file = 'index.html') })
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(crosstalk)
```

## Background

For this project, I decided to recreate Hans Rosling's population and GDP graph. The graph can be found [here](https://www.youtube.com/watch?v=jbkSRLYSojo&ab_channel=BBC). 

Since the video did not share the dataset, I downloaded another dataset from [Our World in Data](https://ourworldindata.org/grapher/life-expectancy-vs-gdp-per-capita). 

## Data Overview and Wrangling

Although the year starts from 1870, I decided to use data only after 1950 because most countries miss values for population and GDP for years before 1950.


```{r, echo = FALSE}
le <- read.csv("data/life_expectancy_data.csv") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  filter(entity != "World",
         year >= 1950) 
```

There are 166 countries in the dataset, but it misses information about which continent a country belongs to.

I used [this website](https://statisticstimes.com/geography/countries-by-continents.php) as a reference to continent, and manually labeled each country by using `case_when` function, but I should have just used `left_join` for the sake of efficiency...

```{r, include = FALSE}
le %>% distinct(entity) %>% nrow()
```

```{r, echo = FALSE}
le <- le %>% 
  mutate(continent = case_when
         (entity %in% c("Algeria", "Angola", "Benin", "Botswana",
                        "Burkina Faso", "Burundi", "Cameroon", "Cape Verde",
                        "Central African Republic", "Chad", "Comoros", "Congo",
                        "Cote d'Ivoire", "Democratic Republic of Congo",
                        "Djibouti", "Egypt", "Equatorial Guinea", "Eswatini",
                        "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea",
                        "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Libya",
                        "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", 
                        "Morocco", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda",
                        "Sao Tome and Principe", "Senegal", "Seychelles",
                        "Sierra Leone", "South Africa", "Sudan", "Tanzania", "Togo",
                        "Tunisia", "Uganda", "Zambia", "Zimbabwe"
                        )
           ~ "Africa",
           
          entity %in% c("Barbados", "Canada", "Costa Rica", "Cuba", "Dominica",
                        "Dominican Republic", "El Salvador", "Guatemala", "Haiti",
                        "Honduras", "Jamaica", "Mexico", "Nicaragua",
                        "Panama", "Puerto Rico", "Saint Lucia", 
                        "Trinidad and Tobago", "United States"
                        ) 
           ~ "North America",
           
          entity %in% c("Argentina", "Bolivia", "Brazil", "Chile", "Colombia",
                        "Ecuador", "Paraguay", "Peru", "Uruguay", "Venezuela"
                        ) 
           ~ "South America",
           
          entity %in% c("Afghanistan", "Azerbaijan", "Bahrain", "Bangladesh",
                        "Cambodia", "China", "Cyprus", "Georgia", "Hong Kong",
                        "India", "Indonesia", "Iran", "Iraq", "Israel", "Japan",
                        "Jordan", "Kazakhstan", "Kuwait", "Kyrgyzstan", "Laos",
                        "Lebanon", "Malaysia", "Moldova", "Mongolia", "Myanmar",
                        "Nepal", "North Korea", "Oman", "Pakistan", "Palestine",
                        "Philippines", "Qatar", "Saudi Arabia", "Singapore",
                        "South Korea", "Sri Lanka", "Syria", "Taiwan", "Tajikistan",
                        "Thailand", "Turkey", "Turkmenistan", "United Arab Emirates",
                        "Uzbekistan", "Vietnam", "Yemen"
                        ) 
           ~ "Asia",
           
          entity %in% c("Albania", "Armenia", "Austria", "Belarus", "Belgium",
                        "Bosnia and Herzegovina", "Bulgaria", "Croatia", "Czechia",
                        "Denmark", "Estonia", "Finland", "France", "Germany", 
                        "Greece", "Iceland", "Ireland", "Italy", "Latvia",
                        "Lithuania", "Luxembourg", "Malta", "Montenegro",
                        "Netherlands", "North Macedonia", "Norway", "Poland",
                        "Portugal", "Romania", "Serbia", "Slovakia", "Slovenia",
                        "Spain", "Sweden", "Switzerland", "Ukraine", 
                        "United Kingdom", "Hungary"
                        ) 
           ~ "Europe",
           
          entity %in% c("Australia", "New Zealand"
                        ) 
           ~ "Oceania"
  ))
```

## Data Visualization

I used plotly and crosstalk to make a visualization.  I think crosstalk does not go well with animation plots as the continents/countries shown are different from ones selected.

```{r message=FALSE, warning=FALSE, echo = FALSE}
shared_le <- SharedData$new(le)

bscols(widths = c(2,NA),
  list(
    filter_select("continent", "Continent", shared_le, ~continent)
  ),
  ggplotly(shared_le %>% 
  ggplot(aes(gdp_per_capita, life_expectancy)) +
  geom_point(aes(size = population_historical_estimates, frame = year, ids = entity, color = continent)) +
  scale_x_log10() +
  labs(x = "GDP per Capita", y = "Life Expectancy", size = ""))
)
```

## Show countries in North America

```{r message=FALSE, warning=FALSE, echo = FALSE}
le_na <- le %>% filter(continent == "North America")
shared_le_na <- SharedData$new(le_na)

bscols(widths = c(2,NA),
  list(
    filter_select("entity", "Country", shared_le_na, ~entity)
  ),
  ggplotly(
  ggplot(data = shared_le_na, aes(gdp_per_capita, life_expectancy)) +
  geom_point(aes(size = population_historical_estimates, frame = year, color = entity)) +
  scale_x_log10() +
  labs(x = "GDP per Capita", y = "Life Expectancy", color = "Country", size = ""))
)
```

## Better One

```{r message=FALSE, warning=FALSE, echo = FALSE}
bscols(widths = c(2,4,6),
  list(
    filter_select("entity", "Country", shared_le, ~entity),
    filter_slider("year", "Years", shared_le, ~year, round = TRUE)
  ),
  add_lines(shared_le %>%
  plot_ly(x = ~year, y = ~life_expectancy, color = ~entity, id = ~entity), showlegend = FALSE) %>% 
    layout(xaxis = list(title = "Year"), yaxis = list(title = "Life Expectancy")),
  
  add_lines(shared_le %>%
  plot_ly(x = ~year, y = ~gdp_per_capita, color = ~entity, id = ~entity) %>% 
    layout(xaxis = list(title = "Year"), yaxis = list(title = "GDP Per Capita")))
)
```

## Improvement

For the improvement, I would like to make a similar graph to the Figure 13.10 from Wilke's [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/time-series.html).

The code was taken and modified from [Wilke's github repo](https://github.com/clauswilke/dataviz/blob/master/time_series.Rmd).

The graph below shows how life expectancy and GDP per capita transition over time in Afghanistan. The transparency of the line represents year. In the future, I would like to incorporate the graph with crosstalk. It seems like crosstalk does not go well with `geom_text_repel`. I also need to figure out how to show only selected country in the graph (in the previous graphs, `color = entity` does it for me, but the graph below has nothing to differentiate countries if that makes sense).

```{r message=FALSE, warning=FALSE, echo = FALSE}
le2 <- le %>% 
  mutate(label = case_when
         (year == "1950" ~ "1950",
          year == "1960" ~ "1960",
          year == "1970" ~ "1970",
          year == "1980" ~ "1980",
          year == "1990" ~ "1990",
          year == "2000" ~ "2000",
          year == "2010" ~ "2010",
          year == "2018" ~ "2018"))
```

```{r message=FALSE, warning=FALSE, echo = FALSE}
library(ggridges)
library(lubridate)
library(ggrepel)

time_series <- le2 %>%
  filter(entity == "Afghanistan") %>% 
  ggplot() +
    aes(gdp_per_capita, life_expectancy, colour = as.numeric(year)) + 
    geom_path(size = 1, lineend = "round") +
  geom_text_repel(
    aes(label = label), point.padding = .2, color = "black",
    min.segment.length = 0, size = 12/.pt,
    nudge_x = le2$nudge_x,
    nudge_y = le2$nudge_y,
    direction = "y"
  ) + 
  scale_colour_gradient(low = "#E7F0FF", high = "#035B8F") + #"#0072b2") +
  guides(colour = FALSE) +
  coord_cartesian(clip = "off") +
  theme(
    axis.ticks.length = unit(0, "pt"),
    plot.margin = margin(21, 14, 3.5, 1.5)) +
  labs(title = "Life Expectancy and GDP change over time in Afghanistan", y = "Life Expectancy", x = "GDP per Capita") +
  theme_bw()

time_series
```

## Reflection

* I wanted to make it as simple as possible. Without the selector for continents/countries, all dots were crammed together and it was difficult to see specific continents/countries. After implementing the selector, I think the visualization became much simpler.

* For the animation visualizations, selected countries do not match with what are shown in the graph, which I want to fix in the future. For the bar graph, the adjacent countries in the list have similar color, so it is hard to distinguish.

* The most frustrating part was definitely the mismatching between selected and shown.
