---
title: "Here's Johnny"
author: "John Calvin"
date: "Spring 2022"
output: html_document
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      output_file = 'index.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
```

## Background

For this project, I decided to recreate Hans Rosling's population and GDP graph. The graph can be found [here](https://www.youtube.com/watch?v=jbkSRLYSojo&ab_channel=BBC). 

Since the video did not share the dataset, I downloaded another dataset from [Our World in Data](https://ourworldindata.org/grapher/life-expectancy-vs-gdp-per-capita). 

## Data Overview and Wrangling

Although the year starts from 1870, I decided to use data only after 1950 because most countries miss values for population and GDP for years before 1950.


```{r}
le <- read.csv("data/life_expectancy_data.csv") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  filter(entity != "World",
         year >= 1950) 
```

```{r}
le %>% head()
```

There are 166 countries in the dataset, but it misses information about which continent a country belongs to.

I used [this website](https://statisticstimes.com/geography/countries-by-continents.php) as a reference to continent, and manually labeled each country.

```{r}
le %>% distinct(entity) %>% nrow()
```

```{r}
le <- le %>% 
  mutate(continent = case_when
         (entity %in% c("Algeria", "Angola", "Benin", "Botswana",
                        "Burkina Faso", "Burundi", "Cameroon", "Cape Verde",
                        "Central African Republic", "Chad", "Comoros", "Congo",
                        "Cote d'Ivoire", "Democratic Republic of Congo",
                        "Djibouti", "Egypt", "Equatorial Guinea", "Eswatini",
                        "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea",
                        "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Libya",
                        "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", 
                        "Morocco", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda",
                        "Sao Tome and Principe", "Senegal", "Seychelles",
                        "Sierra Leone", "South Africa", "Sudan", "Tanzania", "Togo",
                        "Tunisia", "Uganda", "Zambia", "Zimbabwe"
                        )
           ~ "Africa",
           
          entity %in% c("Barbados", "Canada", "Costa Rica", "Cuba", "Dominica",
                        "Dominican Republic", "El Salvador", "Guatemala", "Haiti",
                        "Honduras", "Hungary", "Jamaica", "Mexico", "Nicaragua",
                        "Panama", "Puerto Rico", "Saint Lucia", 
                        "Trinidad and Tobago", "United States"
                        ) 
           ~ "North America",
           
          entity %in% c("Argentina", "Bolivia", "Brazil", "Chile", "Colombia",
                        "Ecuador", "Paraguay", "Peru", "Uruguay", "Venezuela"
                        ) 
           ~ "South America",
           
          entity %in% c("Afghanistan", "Azerbaijan", "Bahrain", "Bangladesh",
                        "Cambodia", "China", "Cyprus", "Georgia", "Hong Kong",
                        "India", "Indonesia", "Iran", "Iraq", "Israel", "Japan",
                        "Jordan", "Kazakhstan", "Kuwait", "Kyrgyzstan", "Laos",
                        "Lebanon", "Malaysia", "Moldova", "Mongolia", "Myanmar",
                        "Nepal", "North Korea", "Oman", "Pakistan", "Palestine",
                        "Philippines", "Qatar", "Saudi Arabia", "Singapore",
                        "South Korea", "Sri Lanka", "Syria", "Taiwan", "Tajikistan",
                        "Thailand", "Turkey", "Turkmenistan", "United Arab Emirates",
                        "Uzbekistan", "Vietnam", "Yemen"
                        ) 
           ~ "Asia",
           
          entity %in% c("Albania", "Armenia", "Austria", "Belarus", "Belgium",
                        "Bosnia and Herzegovina", "Bulgaria", "Croatia", "Czechia",
                        "Denmark", "Estonia", "Finland", "France", "Germany", 
                        "Greece", "Iceland", "Ireland", "Italy", "Latvia",
                        "Lithuania", "Luxembourg", "Malta", "Montenegro",
                        "Netherlands", "North Macedonia", "Norway", "Poland",
                        "Portugal", "Romania", "Serbia", "Slovakia", "Slovenia",
                        "Spain", "Sweden", "Switzerland", "Ukraine", 
                        "United Kingdom"
                        ) 
           ~ "Europe",
           
          entity %in% c("Australia", "New Zealand"
                        ) 
           ~ "Oceania"
  ))
```

## Data Visualization

I found `ppglotly` very useful as it converts ggplot2 to plotly.

```{r}
gg <- le %>% 
  ggplot(aes(gdp_per_capita, life_expectancy)) +
  geom_point(aes(size = population_historical_estimates, frame = year, ids = entity, color = continent)) +
  scale_x_log10() +
  labs(x = "GDP per Capita", y = "Life Expectancy")
ggplotly(gg)
```

I thought this graph was somewhat hard to interpret. We can see the general trend but it is hard to compare different countries. 
To solve this issue, I made a simple function that draws the same graph with specif countries a user chooses.

I am trying to figure out how to make a user interface to make the most of the function I made.

```{r}
make_graph <- function(...) {
  gg1 <- le %>% filter(entity %in% c(...)) %>% 
    ggplot(aes(gdp_per_capita, life_expectancy)) +
    geom_point(aes(size = population_historical_estimates, frame = year, color = entity)) +
    scale_x_log10() +
    labs(x = "GDP per Capita", y = "Life Expectancy")
  ggplotly(gg1)
}

make_graph("United States", "China", "Australia", "Brazil", "Ghana", "Germany")
```



## Reflection
