---
title: "plotly Projects"
author: "Data 303"
date: "Spring 2022"
output: html_document

knit: (
  function(inputFile, encoding) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      output_file = 'index.html') })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The projects

```{r echo = FALSE, results = "asis"}
projects <- dir(pattern = "index.Rmd", recursive = TRUE)
for (p in projects) {
  yaml <- rmarkdown::yaml_front_matter(p)
  glue::glue("* [{title}]({path}) by {name}", 
             path = dirname(p), 
             title = yaml[["title"]],
             name = yaml[["author"]]
  ) |> cat()
}

```

## Instructions

Your task is to create a visualization using plotly in R. 

1. The data you use for this project may not come from an R package, unless it is one that you created.
Otherwise, you may use any data you like. More interesting data makes for more interesting projects.

2. Prepare an Rmd file that produces an `index.html` file. 

    a. The name of your file must end `index.Rmd` 
    
        Take a look at John Calvin's example if you want to 
    name your Rmd something other than `index.Rmd`.
    
    b. The YAML header must include the following items: `title`, `author`
        

3. Prepare an HTML file named **index.html** (generated from your Rmd file).

    The file should contain:
    
    a. Your visualization(s).
    b. Any background information needed to understand the data or the visualization.
    c. A description of the data source with links if the data are available online.
    d. Any other references/citations as needed.
    

4. Your visualization(s) should include at least one interactive element.

5. Also submit your source files. 

    a. You don't need to submit the data if it is available online. (But then your project should also
    load from that online source.)
    
    b. Your project should be buildable by someone who has access to the files you submit. It should not
    rely on other information -- like files on your computer that are not submitted. 
    [Exception: if there is some reason that you cannot share your data, or 
    if your data is larger than 10 MB, talk to me.]

6. Submit your work via a pull request.

    a. Fork this repo: <https://github.com/rpruim/ds303-s22-projects>
    b. Inside `docs/plotly-projects/` create a folder using the naming convention `first-last` (all lower case, no spaces, use a dash for a separator).
    c. Place all your files in this folder.
    d. When you are ready to submit your work, create a pull request. You can find out more about 
    pull requests [here](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/creating-a-pull-request-from-a-fork)
    
